Constraining Supermassive Black Hole Environments.

### 2014-08-04 ms.tex
Fixing some typos

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$>$}}}
\newcommand{\rs}{r_s}
\newcommand{\rb}{r_b}
\newcommand{\vw}{v_w}
\newcommand{\cs}{c_s}

@@

\author[Generozov, Metzger, \& Stone]{Aleksey Generozov$\thanks{E-mail: ag@astro.columbia.edu}$, Nicholas Stone, Brian~D.~Metzger$\footnotemark[1]$\\
Columbia Astropysics Labratory, Columbia University, 550 West 120th Street, New York, NY 10027}

@@

Previous studies used Chandra observations to show that the Bondi accretion power well-correlated with jet power \citep{AllenDunn+:2006a,Fujita+:2013a}. In these studies temperature, and density information are inferred from Chandra x-ray observations, while the jet power is inferred from the energy required to inflate observed bubbles within the x-ray emitting gas.

@@

%In this study we perform hydrodynamic modeling to solve for the steady-state gas density, temperature and velocity profiles near the SMBH sphere of influence, for a sample of nearby, quiescent galaxies. We a adopt the model used by \cite{Quataert:2004a} for the galactic center: gas is supplied by stellar winds, which collide, shock heat, and thermalize.  
We take the same approach. To the best of our knowledge a systematic study of the parameter space of central galaxy properties. Thus, we compute central density profiles for a broad sample of galaxies for \cite{WangMerritt:2004a}--henceforward \citetalias{WangMerritt:2004a}.  This allows us to determine to relate the central stellar properties (e.g. stellar density slope) to the central gas properties.

@@

These or similar equations have been solve many times in the literature (e.g. \citealt{HolzerAxelford1971a,Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}). \citealt{ShcherbakovWong+:2014a} solves similar equation, with additional terms to account for thermal conduction and radiative cooling. However, we find that radiative cooling may be neglected for a large portion of our parameter space--\S~\ref{sec:cooling}.  Additionally thermal conduction may be cut off by magnetic field lines--\S~\ref{sec:conduction}

@@

\begin{align}
\dotMb=4\pi \lambda r_b^2 \rho(r_b) v_{\rm ff}(r_b)
\end{align}

where $rb=G M/\cs^2$.

@@

Where have set $n~1$.  At $\rs$ we have

@@

Thus, $\cs(rs)\simeq \vw(\rs)$. This means that $\dot{M}$ in our model is quite close to the

@@

As noted in \citealt{ShcherbakovWong+:2014a}, it is not obvious whether or not the SNe energy injection treated as a constant throughout time and space, since the flow time-scale may be short compared to the time between successive SNe Ia...
%Thornton and Benton...
\item \emph{MS Pulsars} 
\end{enumerate}

\subsection{Mass accretion rate}
For each of our solutions, we may calculate the accretion rate $\dot{M}$ onto the SMBH. We show $\Medd$ vs. $\Mbh$ as a function of black hole mass for our three  chosen values of the $\vwO$ (200 km/s, 500 km/s, and 1000 km/s). The results are shown in Figure~\ref{fig:mdot_mass}

We find there is a trend of $\eddr$ increasing towards higher black hole masses. However, previous work  suggests a downsizing trend: for quiescent galaxies $L_X \sim \Mbh^\alpha$, where $\alpha\simeq 0.7-0.8$ \citep{Miller+:2013a}. If $L_X\sim\dot{M}^2$, this would suggest $\eddr\sim \Mbh^{-0.6}$. In contrast for fixed $\vwO$ and $\eta$ we obtain that $\eddr\sim \Mbh^{0.5-1}$. One possible solution is that $\eta$ decreases with $\Mbh$. $\dot{M}$ scales linearly with $\eta$, so we would need to have $\eta\sim \Mbh^{-1}$ to explain the observed downsizing trend.

@@

We may compare our results for the $T$ and $\rho$ profiles of the overlapping galaxies, with those inferred from observations.  Note that we we have two free parameters ($\eta$ and $\vwO$ in our model and thus we can can always match $T$ and $\rho$ at least one point.  We generally find that $\vwO=500$ km/s (as would be expected from SNe Ia) gives rough agreement with the observed temperatures.  Thus, we choose $\vwO=500$ km/s.  We choose an $\eta$ so that our solution would be consistent with the observed density profiles. The comparison is shown in Figure~\ref{fig:allen_compare}.

\begin{itemize}
\item \emph{NGC4486} On scales of $\sim1$ kpc the density profile in \citealt{AllenDunn+:2006a}  is shallower than in our model. Our model has a break in the gas density at $\sim 300$ pc--near the location of the Nuker break radius for this galaxy at 560 pc. There is an observed break in the density, but on a scale of a few kpc.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{T_compare.eps}
\includegraphics[width=\columnwidth]{dens_compare.eps}
\caption{\label{fig:allen_compare} Comparison to \citealt{AllenDunn+:2006a}}
\end{figure*}

>>>>

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
        \raise1pt\hbox{$>$}}}
\newcommand{\rs}{r_s}
\newcommand{\rb}{r_b}
\newcommand{\vw}{v_w}

@@

\author[Generozov, Metzger, \& Stone]{Aleksey Generozov$\thanks{E-mail: ag@astro.columbia.edu}$, Brian~D.~Metzger, Nicholas Stone$\footnotemark[1]$\\
Columbia Astropysics Labratory, Columbia University, 550 West 120th Street, New York, NY 10027}

@@

Previous studies used Chandra observations to show that the Bondi accretion power well-correlated with jet power \citep{AllenDunn+:2006a,FujitaKawakatu+:2014a}. In these studies temperature, and density information are inferred from Chandra x-ray observations, while the jet power is inferred from the energy required to inflate observed bubbles within the x-ray emitting gas.

@@

%In this study we perform hydrodynamic modeling to solve for the steady-state gas density, temperature and velocity profiles near the SMBH sphere of influence, for a sample of nearby, quiescent galaxies. We a adopt the model used by \cite{Quataert:2004a} for the galactic center: gas is supplied by stellar winds, which collide, shock heat, and thermalize.  
We take the same approach. To the best of our knowledge a systematic study of the parameter space of central galaxy properties. Thus, we compute central density profiles for a broad sample of galaxies from \citealt{WangMerritt:2004a}--henceforward \citetalias{WangMerritt:2004a}.  This allows us to relate the central stellar properties (e.g. the slope of the stellar density) to the central gas properties.

@@

These or similar equations have been solve many times in the literature (e.g. \citealt{HolzerAxford:1970a,Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}). \citealt{ShcherbakovWong+:2014a} adds additional terms to account for thermal conduction and radiative cooling. However, we find that radiative cooling may be neglected for a large portion of our parameter space--\S~\ref{sec:cooling}.  Additionally thermal conduction may be cut off by magnetic field lines--\S~\ref{sec:conduction}

@@

\begin{align}
\Mdotb=4\pi \lambda r_b^2 \rho(r_b) v_{\rm ff}(r_b)
\end{align}

where $r_b=G M/c_s^2$.

@@

Where have set $n=1$.  At $\rs$ we have

@@

Thus, $\cs(rs)\simeq \vw(\rs)$. This means that $\dot{M}$ in our model is quite close to \Mdotb

@@

As noted in \citealt{ShcherbakovWong+:2014a}, it is not obvious whether or not the SNe energy injection treated as a constant throughout time and space, since the flow time-scale may be short compared to the time between successive SNe Ia...
%AG-Thornton and http://arxiv.org/pdf/1402.6695v2.pdf...
%AG-Case of no stagnation radius
\item \emph{MS Pulsars} 
\end{enumerate}

\subsection{Mass accretion rate}
For each of our solutions, we may calculate the accretion rate $\dot{M}$ onto the SMBH. We show $\Medd$ vs. $\Mbh$ as a function of black hole mass for our three  chosen values of the $\vwO$ (200 km/s, 500 km/s, and 1000 km/s). The results are shown in Figure~\ref{fig:mdot_mass}. 
%%AG-Also want to go into the case of no stagnation radius

We find there is a trend of $\eddr$ increasing towards higher black hole masses. However, previous work  suggests a downsizing trend: for quiescent galaxies $L_X \sim \Mbh^\alpha$, where $\alpha\simeq 0.7-0.8$ \citep{MillerGallo+:2014a}. If $L_X\sim\dot{M}^2$, this would suggest $\eddr\sim \Mbh^{-0.6}$. In contrast for fixed $\vwO$ and $\eta$ we obtain that $\eddr\sim \Mbh^{0.5-1}$. One possible solution is that $\eta$ decreases with $\Mbh$. $\dot{M}$ scales linearly with $\eta$, so we would need to have $\eta\sim \Mbh^{-1}$ to explain the observed downsizing trend.

@@

We may compare our results for the $T$ and $\rho$ profiles with those inferred from observations for the overlapping galaxies.  Note that we we have two free parameters ($\eta$ and $\vwO$ in our model and thus we can can always match $T$ and $\rho$ at least one point.  We generally find that $\vwO=500$ km/s (as would be expected from SNe Ia) gives rough agreement with the observed temperatures.  Thus, we choose $\vwO=500$ km/s.  We choose an $\eta$ so that our solution would be consistent with the observed density profiles. The comparison is shown in Figure~\ref{fig:allen_compare}.
%%AG-The chosen properties, in particular eta may be summarized in a table.

\begin{itemize}
\item \emph{NGC4486} %eta=0.2
On scales of $\sim1$ kpc the density profile in \citealt{AllenDunn+:2006a}  is shallower than in our model. Our model has a break in the gas density at $\sim 300$ pc--near the location of the Nuker break radius for this galaxy at 560 pc. There is an observed break in the density, but on a scale of a few kpc.

@@

\begin{figure*}
\includegraphics[width=\columnwidth]{T_compare.eps}
\includegraphics[width=\columnwidth]{dens_compare.eps}
\caption{\label{fig:allen_compare} Comparison of $n_e$ and $T$ to the \citealt{AllenDunn+:2006a}}. 
%%AG in the future we will have to be carfeful of the mean molecular weight.
\end{figure*}
